---
title: "CME Seat and Commodities Price Forecasting"
subtitle: "ADSP 31006"
author: "Robin Lee"
date: "2025-02-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(zoo)
require(data.table)
require(forecast)
require(tseries)
require(stargazer)
require(DataCombine)
require(knitr)
library(kableExtra)
require(tibble)
library(xts)
library(readxl)
library(glue)
library(dplyr)
library(ggplot2)
require(fitdistrplus)
library(lmtest)
library("TSA")
library(imputeTS)
library(glue)
library(Metrics)
library(corrplot)
library(fGarch)
library(rugarch)
library(bsts)
library(lubridate)
library(tidyr)
```

# Forecasting CME Seat Prices 

## Initialize CME Data

There are three classes of seats CME, IMM, and IOM which confer the rights to trade different sets of commodities traded. CME seat owners can trade everything, IMM seat owners are allowed to trade everything except agricultural products, and IOM seat owners are allowed to trade only in index products and all options.

The seat price files are marked as `***S`. The files contain the price for which CME seats sold and the date of the sale. As expected, the seat price time series is irregularly spaced in time. The three files contain seat prices from the Chicago Mercantile Exchange.

```{r, echo=FALSE, warning=FALSE}
cmes <- read.csv("cmeS.csv")
imms <- read.csv("immS.csv")
ioms <- read.csv("iomS.csv")
kable(head(cmes,4),caption = "Sample of `cmes` Data")
kable(head(imms,4),caption = "Sample of `imms` Data")
kable(head(ioms,4),caption = "Sample of `ioms` Data")
```

\newpage

**Data Interpolation**

\bigskip

To create a time series that contains a seat price for each month starting from January 2001 to December 2013, I use Spline Interpolation because it is specifically geared towards  imputing data with trend.

* This method works by extracting the month and year from the `DateOfSale` variable, and groups the seat price by each month. Because I seek to create a time series that contains a seat price for each month starting from January 2001 to December 2013, I average each month's price if there exists more than one. Then, I create a filler data frame with all the months from 2001 to 2013, and merge the seat data frame with the month filler. 

* By doing so, I have the full month and year range in the seat data, with NA values for months without any prices. Finally, I use spline interpolation to fill any NA values with an estimated mean price. Spline interpolation can create continuous curves from discrete data points, which makes it a bit more realistic than a simple linear interpolation.

\bigskip

```{r, echo=F, warning=FALSE}
interpolate_ts <- function(df, class, show_plot=FALSE){
  df$DateOfSale <- as.Date(df$DateOfSale, format = "%m/%d/%Y")
  df$YearMonth <- as.Date(paste0(df$Year, "-", df$Month, "-01"), format = "%Y-%m-%d")
  df_yearmonth <- df %>% group_by(YearMonth) %>% 
    summarise(mean_price = mean(price,na.rm=T)) 
  
  all_months <- data.frame(YearMonth = seq(from = as.Date("2001-01-01"), 
                                        to = as.Date("2013-12-01"), 
                                        by = "month"))
  
  df_full <- merge(all_months, df_yearmonth, by = "YearMonth", all.x = TRUE)

  ## Backward Fill
  # df_full$mean_price_filled <- na_locf(df_full$mean_price, option = "locf", 
  #                                       na_remaining = "rev") 
  ## KNN Imputation
  # library(DMwR)
  # df_full$mean_price_filled <- knnImputation(df_full$mean_price, k = 5)
  
  ## Spline Interpolation
  df_full$mean_price_filled <- na.spline(df_full$mean_price)
  
  if (show_plot == TRUE){
    ## Code for plotting:
    # Only keep the interpolated prices
    df_full$only_filled <- ifelse(is.na(df_full$mean_price), 
                                  df_full$mean_price_filled, NA)
    # Convert to time series format
    df_full_ts <- ts(df_full$mean_price, 
                start = c(2001, 1), frequency = 12)
    ggplot() + 
      geom_line(data=df_full, aes(x=YearMonth,y=mean_price_filled,color="Interpolated Data"),
                size=1) +
      geom_point(data=df_full[!is.na(df_full$only_filled), ], 
                 aes(x=YearMonth,y=only_filled,color="Interpolated Data"),
                 size=2) +
      geom_line(data=df_full, aes(x=YearMonth,y=mean_price,color="Original"),size=1) +
      labs(title = glue("CME Seat Prices Sold - {class}"),x="Year",y="Seat Price") +
      scale_color_manual(values = c("Original" = "black", "Interpolated Data" = "red")) +
      theme_minimal() 
  } else {
    return(df_full)
  }
}
```

```{r, echo=FALSE, warning=FALSE,fig.width=12,fig.height=5}
interpolate_ts(cmes, "CMES", show_plot = T)
cmes_interpolate <- interpolate_ts(cmes, "CMES", show_plot = F)
cmes_interpolate <- cmes_interpolate %>% dplyr::select(-(mean_price)) %>% 
  rename(Date=YearMonth)
```

```{r, echo=FALSE, warning=FALSE,fig.width=12,fig.height=5}
interpolate_ts(imms, "IMMS", show_plot = T)
imms_interpolate <- interpolate_ts(imms, "IMMS", show_plot = F)
imms_interpolate <- imms_interpolate %>% dplyr::select(-(mean_price)) %>% 
  rename(Date=YearMonth)
```

```{r, echo=FALSE, warning=FALSE,fig.width=12,fig.height=5}
interpolate_ts(ioms, "IOMS", show_plot = T)
ioms_interpolate <- interpolate_ts(ioms, "IOMS", show_plot = F)
ioms_interpolate <- ioms_interpolate %>% dplyr::select(-(mean_price)) %>% 
  rename(Date=YearMonth)
```

The file `contracts_volume` contains the monthly trading volume for different CME products. Each row has the first date of the following month (the trade volume is for the month), the commodity indicator for CME, the short description of the commodity, whether it is Future (F) or Option (O), the electronic trading volume and the total trading volume. The floor trading volume is the difference between the total volume and the electronic volume.

```{r, echo=FALSE, warning=FALSE}
contracts_volume <- read.csv("Contracts_Volume.csv")
#contracts_volume
contracts_volume$Date <- as.Date(contracts_volume$Date, format = "%m/%d/%Y")
contracts_volume$Electronic.Volume <- 
  as.numeric(gsub(",","", contracts_volume$Electronic.Volume))
kable(head(contracts_volume,4),caption = "Sample of `contracts_volume` Data")
```

The file `contracts_classification` associates each CME commodity with a seat class.

```{r, echo=FALSE, warning=FALSE}
contracts_classification <- read.csv("Contracts_Classification.csv")
#contracts_classification
kable(head(contracts_classification,4),caption = "Sample of `contracts_classification` Data")
```

Note that owning a CME seat gives a trader advantage for floor trading. In contrast, owning a CME seat does not have any advantage for electronic trading, since electronic trading constitutes market participants trading directly with other participants via CME's network connections and order matching engine.

\bigskip

There are three classes of seats CME, IMM, and IOM which confer the rights to trade different sets of commodities traded. CME seat owners can trade everything, IMM seat owners are allowed to trade all futures and options except agricultural products, and IOM seat owners are allowed to trade future and options only in index products and all other CME options.

To find the volume, for example index products, add both the futures and options volume for the same index product for a month.

\bigskip

**Data Aggregation**

Commodities are traded both on the floor and electronically. Because the volume data set only has total volume and electronic volume, I still need to calculate the floor volume. Thus, I sort out of the volume data, specifically for those commodities that are relevant for the particular badge (CME, IMM, IOM). Then, I aggregate the data for each Commodity Indicator for each month, ignoring the distinctions between futures and options.

```{r,include = FALSE,eval=FALSE}
filter_volume <-
  function(.data,
           .division,
           .date = list(floor = "2001-01-01", ceiling = "2012-12-01"),
           .date_format = "%Y-%m-%d",
           test_set = FALSE) {
    dates <- map(.date, ~ lubridate::as_date(.x, .date_format))
    if (.division %in% c("iom", "imm")) {
      .data <- .data %>% filter(division == .division)
    }
    
    if (.division == "cme") {
      .data <- .data %>% 
        mutate(division = "cme") %>% 
        distinct()
    }
    
    if (test_set) {
      .data <- .data %>% filter(date >= dates)
    } else {
      .data <- .data %>% filter(between(date, dates$floor, dates$ceiling))
    }
    .data
  }
aggregate_volume <- function(.data) {
  .data %>% 
    group_by(date) %>%
    summarize(
      total_volume = sum(total_volume) %>% as.double(),
      electronic_volume = sum(electronic_volume) %>% as.double(),
      floor_volume = sum(floor_volume) %>% as.double()
    ) %>% 
    mutate(year_month = yearmonth(date)) %>% 
    select(year_month, everything(), -date)
}
transform_by_division <- function(.division, .data, ...) {
  .data %>%
    filter_volume(.division, ...) %>%
    aggregate_volume()
}

divisions <-
  c(
    cme = "cme",
    imm = "imm",
    iom = "iom"
  )

contract_volume_divisions <- contract_volume_divisions %>%
  mutate(date = as.Date(date, format="%m/%d/%Y"))

train_volumes <-
  map(divisions, transform_by_division, .data = contract_volume_divisions)

test_volumes <-
  map(
    divisions,
    ~ transform_by_division(
      .division = .x,
      .data = contract_volume_divisions,
      .date = "2013-01-01",
      test_set = TRUE
    )
  )
```

```{r, echo=FALSE}
# calculate the Floor volume:
contracts_volume$Floor.Volume = contracts_volume$Total.Volume - contracts_volume$Electronic.Volume
# add division for each commodity code
contracts <- merge(contracts_volume, contracts_classification, 
                   by.x = "Commodity.Indicator", 
                   by.y = "Commodity.Code", 
                   all.x = TRUE)

# prevent duplicates for CME since CME is all
contracts_classification_distinct <- 
  contracts_classification %>% distinct(Commodity.Code, .keep_all = TRUE)
contracts_distinct <- merge(contracts_volume, contracts_classification_distinct, 
                   by.x = "Commodity.Indicator", 
                   by.y = "Commodity.Code", 
                   all.x = TRUE)

# sort out of the volume data, those commodities that are relevant for the particular badge
cme_volume <- contracts_distinct
imm_volume <- contracts[contracts$Division=="IMM",]
iom_volume <- contracts[contracts$Division=="IOM",]

cme_monthly <- cme_volume %>%
  group_by(Date) %>%
  summarize(
    Elec.Vol = sum(Electronic.Volume, na.rm = TRUE),
    Tot.Vol = sum(Total.Volume, na.rm = TRUE),
    Flo.Vol = sum(Floor.Volume, na.rm = TRUE))
imm_monthly <- imm_volume %>%
  group_by(Date) %>%
  summarize(
    Elec.Vol = sum(Electronic.Volume, na.rm = TRUE),
    Tot.Vol = sum(Total.Volume, na.rm = TRUE),
    Flo.Vol = sum(Floor.Volume, na.rm = TRUE))
iom_monthly <- iom_volume %>%
  group_by(Date) %>%
  summarize(
    Elec.Vol = sum(Electronic.Volume, na.rm = TRUE),
    Tot.Vol = sum(Total.Volume, na.rm = TRUE),
    Flo.Vol = sum(Floor.Volume, na.rm = TRUE))
```

```{r, echo=FALSE, warning=FALSE}
cme_final <- left_join(cmes_interpolate,cme_monthly,by="Date")
kable(head(cme_final,4),caption = "Sample of CME Data")

imm_final <- left_join(imms_interpolate,imm_monthly,by="Date")
kable(head(imm_final,4),caption = "Sample of IMM Data")

iom_final <- left_join(ioms_interpolate,iom_monthly,by="Date")
kable(head(iom_final,4),caption = "Sample of IOM Data")
```

\bigskip

I seek to use the trading volume information to forecast the CME monthly seat price for 2013. Following best practices, I first conduct exploratory data analysis to find initial data relationships, such as correlations. For example, the total trade volume for all CME products might be a good predictor for CME seat class, but not for the others. As such, I choose and select which commodities have the most influence on the IMM and IOM seat prices.

\bigskip

**Exploratory Data Analysis**

\bigskip

```{r, echo=FALSE,fig.width=12,fig.height=5}
par(mfrow = c(1, 3))
plot.ts(cme_final$Date,cme_final$Tot.Vol,type='l',ylab = "Volume",
        xlab="Date",ylim=c(0,max(cme_final$Tot.Vol)),main="CME Volume Breakdown")
lines(cme_final$Date,cme_final$Elec.Vol,col = "red")
lines(cme_final$Date,cme_final$Flo.Vol,col = "blue")
legend("topleft", legend = c("Total","Electric","Floor"), col = c("black","red","blue"), 
       lty = c(2,1,1), lwd = 2)

plot.ts(imm_final$Date,imm_final$Tot.Vol,type='l',ylab = "Volume",
        xlab="Date",ylim=c(0,max(imm_final$Tot.Vol)),main="IMM Volume Breakdown")
lines(imm_final$Date,imm_final$Elec.Vol,col = "red")
lines(imm_final$Date,imm_final$Flo.Vol,col = "blue")
legend("topleft", legend = c("Total","Electric","Floor"), col = c("black","red","blue"), 
       lty = c(2,1,1), lwd = 2)

plot.ts(iom_final$Date,iom_final$Tot.Vol,type='l',ylab = "Volume",
        xlab="Date",ylim=c(0,max(iom_final$Tot.Vol)),main="IOM Volume Breakdown")
lines(iom_final$Date,iom_final$Elec.Vol,col = "red")
lines(iom_final$Date,iom_final$Flo.Vol,col = "blue")
legend("topleft", legend = c("Total","Electric","Floor"), col = c("black","red","blue"), 
       lty = c(2,1,1), lwd = 2)
```

```{r, echo=FALSE,fig.width=12,fig.height=5}
par(mfrow = c(1, 3))
corrplot(cor(cme_final[,-1]), method = 'square',type="upper", addCoef.col = 'black')
title("CME Correlation")
corrplot(cor(imm_final[,-1]), method = 'square',type="upper", addCoef.col = 'black')
title("IMM Correlation")
corrplot(cor(iom_final[,-1]), method = 'square',type="upper", addCoef.col = 'black')
title("IOM Correlation")
```

- Based off the three correlation plots above, we will use `Total Volume` for CME and IMM, as these volumes have the highest correlation. Likewise, we will use `Floor Volume` for IOM, as it is significantly higher than Electric or Total.

## 1. Linear regression 

* Assuming seat price is dependent and volume(s) are independent

```{r,fig.width=12,fig.height=5}
linear_regression <- function(data, label){
  if (label == "IOM"){
    model <- lm(mean_price_filled ~ Flo.Vol, data = data)
  } else {
    model <- lm(mean_price_filled ~ Tot.Vol, data = data)
  }
  return(model)
  }
```

**1.1. CME Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(linear_regression(cme_final,"CME"),test=F)
```

**1.2. IMM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(linear_regression(imm_final,"IMM"),test=F)
```

**1.3. IOM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(linear_regression(iom_final,"IOM"),test=F)
```

## 2. Linear regression with ARMA Errors

```{r}
linear_arma <- function(data,label){
  data_ts <- ts(data$mean_price_filled, start = c(2001, 1), frequency = 12)
  if (label == "IOM"){
    xreg <- ts(data$Flo.Vol, start = c(2001, 1), frequency = 12)
  } else {
    xreg <- ts(data$Tot.Vol, start = c(2001, 1), frequency = 12)
  }
  model <- auto.arima(data_ts, xreg = xreg)
  return(model)
  }
```

**2.1. CME Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(linear_arma(cme_final,"CME"),test=F)
```

**2.2. IMM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(linear_arma(imm_final,"IMM"),test=F)
```

**2.3. IOM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(linear_arma(iom_final,"IOM"),test=F)
```

## 3. Holt Winters

```{r}
holt_winters <- function(data){
  data_ts <- ts(data$mean_price_filled, start = c(2001, 1), frequency = 12)
  model <- HoltWinters(data_ts,seasonal="mult")
  return(model)
  }
```

**3.1. CME Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(holt_winters(cme_final),test=F)
```

**3.2. IMM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(holt_winters(imm_final),test=F)
```

**3.3. IOM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(holt_winters(iom_final),test=F)
```

## 4. ARIMA

```{r}
arima_default <- function(data){
  data_ts <- ts(data$mean_price_filled, start = c(2001, 1), frequency = 12)
  model <- auto.arima(data_ts,seasonal=FALSE)
  return(model)
  }
```

**4.1. CME Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(arima_default(cme_final),test=F)
```

**4.2. IMM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(arima_default(imm_final),test=F)
```

**4.3. IOM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(arima_default(iom_final),test=F)
```

## 5. Seasonal ARIMA (SARIMA)

```{r}
arima_seasonal <- function(data){
  data_ts <- ts(data$mean_price_filled, start = c(2001, 1), frequency = 12)
  model <- auto.arima(data_ts,D=1)
  return(model)
  }
```

**5.1. CME Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(arima_seasonal(cme_final),test=F)
```

**5.2. IMM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(arima_seasonal(imm_final),test=F)
```

**5.3. IOM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(arima_seasonal(iom_final),test=F)
```


## 6. Fractional ARIMA (ARFIMA)

```{r}
arima_fractional <- function(data){
  data_ts <- ts(data$mean_price_filled, start = c(2001, 1), frequency = 12)
  model <- forecast::arfima(data_ts)
  return(model)
  }
```

**6.1. CME Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(arima_fractional(cme_final),test=F)
```

**6.2. IMM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(arima_fractional(imm_final),test=F)
```

**6.3. IOM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(arima_fractional(iom_final),test=F)
```


## 7. ARMA and GARCH combination 

* Using the `fGarch` R library and `garchFit()`:

```{r}
garch <- function(data){
  data$mean_price_filled <- scale(data$mean_price_filled)
  data_ts <- ts(data$mean_price_filled, start = c(2001, 1), frequency = 12)
  arma_model <- auto.arima(data_ts, seasonal = FALSE)
  garch_model <- garchFit(~ garch(1,1), data = residuals(arma_model), trace = FALSE)
  return(garch_model) 
}
```

**7.1. CME Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(residuals(garch(cme_final)),test=F)
```

**7.2. IMM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(residuals(garch(imm_final)),test=F)
```

**7.3. IOM Division**

```{r,fig.width=12,fig.height=5, echo=FALSE}
checkresiduals(residuals(garch(iom_final)),test=F)
```

## 8. Model Comparisons

```{r, echo=FALSE, warning=FALSE}
model_evaluation <- function(data, label){
  train <- data[data$Date < '2013-01-01',]
  test <- data[data$Date >= '2013-01-01',]
  compute_rmse <- function(actual, predicted) {
    return(rmse(actual, predicted))
  }
  pred_linear <- predict(linear_regression(train, "CME"), test)
  if (label == "IOM"){
    xreg <- ts(train$Flo.Vol, start = c(2001, 1), frequency = 12)
  } else {
    xreg <- ts(train$Tot.Vol, start = c(2001, 1), frequency = 12)
  }
  pred_arma <- forecast(linear_arma(train, "CME"), xreg = xreg)$mean
  pred_hw <- forecast(holt_winters(train), h = nrow(test))$mean
  pred_arima_def <- forecast(arima_default(train), h = nrow(test))$mean
  pred_arima_seas <- forecast(arima_seasonal(train), h = nrow(test))$mean
  pred_arima_frac <- forecast(arima_fractional(train), h = nrow(test))$mean
  pred_garch <- predict(garch(train), n.ahead = nrow(test))$meanForecast
  # Compute RMSE for each model
  rmse_linear <- compute_rmse(test$mean_price_filled, pred_linear)
  rmse_arma <- compute_rmse(test$mean_price_filled, pred_arma)
  rmse_hw <- compute_rmse(test$mean_price_filled, pred_hw)
  rmse_arima_def <- compute_rmse(test$mean_price_filled, pred_arima_def)
  rmse_arima_seas <- compute_rmse(test$mean_price_filled, pred_arima_seas)
  rmse_arima_frac <- compute_rmse(test$mean_price_filled, pred_arima_frac)
  rmse_garch <- compute_rmse(test$mean_price_filled, pred_garch)
  rmse_results <- data.frame(
    Model = c("Linear Regression", "ARMA", "Holt-Winters", "ARIMA Default", 
              "ARIMA Seasonal", "ARIMA Fractional", "GARCH"),
    RMSE = c(rmse_linear, rmse_arma, rmse_hw, rmse_arima_def, 
             rmse_arima_seas, rmse_arima_frac, rmse_garch)
  )
  return(rmse_results)
}
```

**8.1. CME Division**
```{r, echo=FALSE, warning=FALSE}
kable(model_evaluation(cme_final,"CME"))
```

**8.2. IMM Division**
```{r,warning=FALSE, echo=FALSE, warning=FALSE}
kable(model_evaluation(imm_final,"IMM"))
```

**8.3. IOM Division**
```{r, echo=FALSE, warning=FALSE}
kable(model_evaluation(iom_final,"IOM"))
```

\newpage

# Predicting Commodity Prices using BSTS

`sts_data_2023` is quarterly data for 2 commodities (CL & GC), 1 macroeconomic indicator (CPI) and 1 Index (SPX).

```{r,echo=FALSE, warning=FALSE}
sts_data_2023 <- read.csv("sts_data_2023.csv")
sts_data_2023$Date <- as.Date(sts_data_2023$Date, format = "%m/%d/%y")
sts_data_2023$Date <- as.Date(paste0(ifelse(as.numeric(format(sts_data_2023$Date, "%Y")) > 2024, 
                               as.numeric(format(sts_data_2023$Date, "%Y")) - 100, 
                               format(sts_data_2023$Date, "%Y")), 
                        "-", format(sts_data_2023$Date, "%m-%d")))
sts_data_2023 <- na.omit(sts_data_2023)
kable(head(sts_data_2023,4),caption = "Sample of `sts_data_2023` Data")
```

```{r,fig.width=12,fig.height=5,echo=FALSE}
par(mfrow = c(1, 2))
plot.ts(sts_data_2023$Mean_CL, main = "Mean CL Over Time",ylab = "Mean CL")
plot.ts(sts_data_2023$Mean_GC, main = "Mean GC Over Time",ylab = "Mean GC")
```

In this section, I use the BSTS package to fit models and predict CL and GC prices for 5 periods after:

-   6/1979 (fit1)
-   6/2004 (fit2)
-   12/2022 (fit3)

The models for each CL and GC use different State specifications with:

1.  Dynamic regression of other 3 variables (lags)
2.  Different trend component for each model

\bigskip

```{r,message=TRUE, warning=FALSE}
#data <- sts_data_2023
#date1<- "6/1979"
bsts_forecasts <- function(data, date1){
  fit_end <- ceiling_date(as.Date(paste0("01/", date1), format="%d/%m/%Y"),"month") 
  data_fit <- data %>% filter(Date <= fit_end)
  
  # CL: Define state space specification
  data_fit_cl <- data_fit %>% 
    mutate(
      GC_Lag1 = lag(Mean_GC, 1),
      CPI_Lag1 = lag(Mean_CPI, 1),
      SPX_Lag1 = lag(Mean_SPX, 1)
      ) %>%
    drop_na() # Remove NA values after lagging
  
  ss1_cl <- list()
  ss1_cl <- AddLocalLevel(ss1_cl, y = data_fit_cl$Mean_CL)
  
  ss2_cl <- list()
  ss2_cl <- AddLocalLinearTrend(ss2_cl, y = data_fit_cl$Mean_CL)
  
  ss3_cl <- list()
  ss3_cl <- AddLocalLinearTrend(ss3_cl, y = data_fit_cl$Mean_CL) 
  ss3_cl <- AddSeasonal(ss3_cl, y = data_fit_cl$Mean_CL, nseasons = 4)
  
  sink("/dev/null")
  # Fit BSTS model with lags 
  model1_cl <- bsts(
    Mean_CL ~ GC_Lag1 + CPI_Lag1 + SPX_Lag1, 
    state.specification = ss1_cl,
    data = data_fit_cl,
    niter = 1000
    )
  model2_cl <- bsts(
    Mean_CL ~ GC_Lag1 + CPI_Lag1 + SPX_Lag1, 
    state.specification = ss2_cl,
    data = data_fit_cl,
    niter = 1000
    )
  model3_cl <- bsts(
    Mean_CL ~ GC_Lag1 + CPI_Lag1 + SPX_Lag1,
    state.specification = ss3_cl,
    data = data_fit_cl,
    niter = 1000
    )
  sink()
  
  
  # GC: Define state space specification
  data_fit_gc <- data_fit %>%
    mutate(
      CL_Lag1 = lag(Mean_CL, 1), 
      CPI_Lag1 = lag(Mean_CPI, 1), 
      SPX_Lag1 = lag(Mean_SPX, 1)
      ) %>%
    drop_na() # Remove NA values after lagging
  
  ss1_gc <- list()
  ss1_gc <- AddLocalLevel(ss1_gc, y = data_fit_gc$Mean_GC)
  
  ss2_gc <- list()
  ss2_gc <- AddLocalLinearTrend(ss2_gc, y = data_fit_gc$Mean_GC)
  
  ss3_gc <- list()
  ss3_gc <- AddLocalLinearTrend(ss3_gc, y = data_fit_gc$Mean_GC) 
  ss3_gc <- AddSeasonal(ss3_gc, y = data_fit_gc$Mean_GC, nseasons = 4)
  
  # Fit BSTS model with lags
  sink("/dev/null")
  model1_gc <- bsts(
    Mean_GC ~ CL_Lag1 + CPI_Lag1 + SPX_Lag1,
    state.specification = ss1_gc,
    data = data_fit_gc,
    niter = 1000
    )
  model2_gc <- bsts(
    Mean_GC ~ CL_Lag1 + CPI_Lag1 + SPX_Lag1,
    state.specification = ss2_gc,
    data = data_fit_gc,
    niter = 1000
    )
  model3_gc <- bsts(
    Mean_GC ~ CL_Lag1 + CPI_Lag1 + SPX_Lag1,
    state.specification = ss3_gc,
    data = data_fit_gc,
    niter = 1000
    ) 
  sink()
  
  horizon <- 5
  
  # CL prediction
  last_values_cl <- tail(data_fit_cl, 1) # Take last available values 
  newdata_cl <- data.frame(
    GC_Lag1 = rep(last_values_cl$GC_Lag1, horizon),
    CPI_Lag1 = rep(last_values_cl$CPI_Lag1, horizon),
    SPX_Lag1 = rep(last_values_cl$SPX_Lag1, horizon)
    )
  # Predict using BSTS
  predictions1_cl <- predict(model1_cl, horizon = horizon, newdata = newdata_cl) 
  predictions2_cl <- predict(model2_cl, horizon = horizon, newdata = newdata_cl) 
  predictions3_cl <- predict(model3_cl, horizon = horizon, newdata = newdata_cl)
  
  actual_cl <- data %>% filter(Date >= fit_end-months(2)) %>% head(5) %>% 
    dplyr::select(Mean_CL)
  
  # GC prediction
  last_values_gc <- tail(data_fit_gc, 1) # Take last available values 
  newdata_gc <- data.frame(
    CL_Lag1 = rep(last_values_gc$CL_Lag1, horizon),
    CPI_Lag1 = rep(last_values_gc$CPI_Lag1, horizon),
    SPX_Lag1 = rep(last_values_gc$SPX_Lag1, horizon)
    )
  # Predict using BSTS
  predictions1_gc <- predict(model1_gc, horizon = horizon, newdata = newdata_gc)
  predictions2_gc <- predict(model2_gc, horizon = horizon, newdata = newdata_gc)
  predictions3_gc <- predict(model3_gc, horizon = horizon, newdata = newdata_gc)
  
  actual_gc <- data %>% filter(Date >= fit_end-months(2)) %>% head(5) %>% 
    dplyr::select(Mean_GC)
  
  if (date1=="12/2022"){
    cat(" RMSE (CL Model 1):",Metrics::rmse(actual_cl$Mean_CL,
                                            predictions1_cl$mean[1:2]),"\n", 
        "RMSE (CL Model 2):",Metrics::rmse(actual_cl$Mean_CL,
                                           predictions2_cl$mean[1:2]),"\n", 
        "RMSE (CL Model 3):",Metrics::rmse(actual_cl$Mean_CL,
                                           predictions3_cl$mean[1:2]),"\n", 
        "RMSE (GC Model 1):",Metrics::rmse(actual_gc$Mean_GC,
                                           predictions1_gc$mean[1:2]),"\n", 
        "RMSE (GC Model 2):",Metrics::rmse(actual_gc$Mean_GC,
                                           predictions2_gc$mean[1:2]),"\n", 
        "RMSE (GC Model 3):",Metrics::rmse(actual_gc$Mean_GC,
                                           predictions3_gc$mean[1:2]))
    } else {
      cat(" RMSE (CL Model 1):",Metrics::rmse(actual_cl$Mean_CL,predictions1_cl$mean),"\n",
          "RMSE (CL Model 2):",Metrics::rmse(actual_cl$Mean_CL,predictions2_cl$mean),"\n", 
          "RMSE (CL Model 3):",Metrics::rmse(actual_cl$Mean_CL,predictions3_cl$mean),"\n", 
          "RMSE (GC Model 1):",Metrics::rmse(actual_gc$Mean_GC,predictions1_gc$mean),"\n", 
          "RMSE (GC Model 2):",Metrics::rmse(actual_gc$Mean_GC,predictions2_gc$mean),"\n", 
          "RMSE (GC Model 3):",Metrics::rmse(actual_gc$Mean_GC,predictions3_gc$mean))
      }
  # Plots
  par(mfrow = c(2, 3))
  plot(predictions1_cl, main = "CL: Dynamic regression of other 3 variables") 
  plot(predictions2_cl, main = "CL: Different trend component for each model") 
  plot(predictions3_cl, main = "CL: State specification of Our Choice (Seasonal)") 
  plot(predictions1_gc, main = "GC: Dynamic regression of other 3 variables") 
  plot(predictions2_gc, main = "GC: Different trend component for each model") 
  plot(predictions3_gc, main = "GC: State specification of Our Choice (Seasonal)")
  
  # Inclusion Probability Plots
  par(mfrow = c(2, 3))
  plot(model1_cl,"coef", main = "CL: Dynamic regression of other 3 variables")
  plot(model2_cl,"coef", main = "CL: Different trend component for each model")
  plot(model3_cl,"coef", main = "CL: State specification of Our Choice (Seasonal)")
  plot(model1_gc,"coef", main = "GC: Dynamic regression of other 3 variables")
  plot(model2_gc,"coef", main = "GC: Different trend component for each model")
  plot(model3_gc,"coef", main = "GC: State specification of Our Choice (Seasonal)")
  mtext(" ", outer = TRUE, cex = 1.5, line = -1)
}
```



## Fit 1: 6/1979

```{r,message=FALSE,fig.width=12,fig.height=6, warning=FALSE}
bsts_forecasts(sts_data_2023,"6/1979")
```

\bigskip

For Fit 1 (6/1979), when predicting CL, we can see that the lags of CPI are useful for Models 2 and 3, while a lag of GC is useful for Model 1. Contrastly, when predicting GC, the lags of CL are useful when using Model 1, while CPI is useful for Model 2. Finally, SPX lag is useful in predicting GC for Model 3.  

## Fit 2: 6/2004

```{r,message=FALSE,fig.width=12,fig.height=6, warning=FALSE}
bsts_forecasts(sts_data_2023,"6/2004")
```

\bigskip

For Fit 2 (6/2004), when predicting CL, we can see that the lags of CPI and GC are useful when using Models 1 and 2, though significantly more for Model 1. It seems like lags are not very useful when using Model 3, though we see a slight increase for SPX lag. Contrastly, when predicting GC, the lags of CL are useful when using Model 1. However, lags of other variables are not very useful in predicting GC when using Model 2 and 3, though both have slight increases with SPX lag.  

## Fit 3: 12/2022

```{r,message=FALSE,fig.width=12,fig.height=6, warning=FALSE}
bsts_forecasts(sts_data_2023,"12/2022")
```

\bigskip

For Fit 3 (12/2022), when predicting CL, we see that Model 1 and Model 2 do not benefit much from lags of other variables, though both suggest CPI may be useful. However, CPI is noticeably more useful for Model 3, as well as a little from GC and SPX. Constrastly, when predicting GC, CPI lag is very useful for Model 1. Lags of other variables are not useful for Model 2, while CL lag is slightly useful in predicting GC when using Model 3. 





















